---
title: "Network Visualization: Relationship of Chemical Mixtures to Brain Metabolite level, Cortical Thickness and Brain Volumes"
author: "Virgilio Gonzenbach, Zheng Ren, Quy Cao"
date: "`r Sys.Date()`"

output: 
  html_document:
    page_layout: full
    toc: true
    toc_title: 'Contents'
    toc_location: left
    toc_float: true
    df_print: paged
    code_folding: hide
    code_overflow: wrap
    theme: journal
    code_tools: true
    fig_width: 10
    fig_height: 7
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      error = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r data_prep}
#| message: false
library(tidyr)
library(dplyr)
library(network)
library(stringr)
library(igraph)
library(visNetwork)

# Read in Brain Metabolite data
nt_df <- read.csv(here::here('data/brain_metabolites_interaction.csv')) %>% 
   tibble::add_column(outcome.type = 'Brain Metabolite', .after  = 'outcome') %>% 
  tibble::add_column(effect = 'interaction', .after  = 'outcome.type')

# Read in cortical thickness data
thick_df <- rbind(
  here::here('data/cortical_thickness_interaction.csv') %>% 
    read.csv() %>%
    tibble::add_column(effect = 'interaction'),
here::here('data/cortical_thickness_main.csv') %>% 
    read.csv() %>%
    tibble::add_column(window = 'all', .after = 'term') %>% 
    tibble::add_column(effect = 'main')
  )  %>% 
  tibble::add_column(outcome.type = 'Cortical Thickness', .after  = 'outcome')

# Read in volumes data
vol_df <- rbind(
  here::here('data/cortical_volumes_interaction.csv') %>% 
    read.csv() %>%
    tibble::add_column(effect = 'interaction'),
  here::here('data/cortical_volumes_main.csv') %>% 
    read.csv() %>%
    tibble::add_column(window = 'all', .after = 'term') %>% 
    tibble::add_column(effect = 'main')
  ) %>% 
  tibble::add_column(outcome.type = 'Volume', .after  = 'outcome')
  
# Read in subcortical volumes data
subcort_df <- rbind(
  here::here('data/subcortical_volumes_interaction.csv') %>% 
    read.csv() %>%
    tibble::add_column(effect = 'interaction'),
  here::here('data/subcortical_volumes_main.csv') %>% 
    read.csv() %>%
    tibble::add_column(window = 'all', .after = 'term') %>% 
    tibble::add_column(effect = 'main')
  ) %>% 
  tibble::add_column(outcome.type = 'Subcortical Volume', .after  = 'outcome')

# Join all data.frames
df <- rbind(nt_df, 
            thick_df, 
            vol_df,
            subcort_df)
```


## Summary 

The present notebook visualizes statistically significant effects of chemical mixtures (OPEs) to different outcomes: brain metabolites, cortical thickness estimates and cortical and subcortical brain volumes. A Multiple Informant Model was applied to test main effects and window specific (interaction) effects according to the following formula: 

> **Outcome ∼ log(OPE) * window + Cotinine + Sex + Race + Home Score + Maternal Education**

```{r}
## create nodes 
nodes <- dplyr::union(df %>% select(term, outcome.type, window) %>% rename(variable=term), 
             df %>% select(outcome, outcome.type, window) %>% rename(variable=outcome)) %>%  
  mutate(outcome.type = ifelse(variable %in% c("BCEtP", "BDCPP", "DBuP", "DPHP"), 'Chem. Mix.', outcome.type)) %>% 
  distinct() %>% 
  mutate(value = 1) %>% 
  pivot_wider(names_from = 'window', values_from = 'value', values_fill = 0) %>% 
  mutate(variable = ifelse(outcome.type == 'Cortical Thickness', paste(variable, '(CT)'), variable)) %>% 
  mutate(variable = ifelse(outcome.type == 'Volume', paste(variable, '(Vol.)'), variable)) %>% 
  mutate(id = seq_len(nrow(.)), 
         .before = 'variable') %>% 
  rename(group = outcome.type)
```

```{r}
id_node <- function(node.name){
  nodes %>% 
    filter(variable == node.name) %>% 
    pull(id)
}

links <- df %>% 
  mutate(outcome = ifelse(outcome.type == 'Cortical Thickness', paste(outcome, '(CT)'), outcome)) %>% 
  mutate(outcome = ifelse(outcome.type == 'Volume', paste(outcome, '(Vol.)'), outcome)) %>% 
  rename(from = term,to = outcome) %>% 
  mutate(to = purrr::map_int(to, id_node),
         from = purrr::map_int(from, id_node), 
         dashes = effect == 'interaction', 
         estimate = ifelse(sign(estimate) == 1, 'positive', 'negative')) %>% 
  select(from, to, window, effect, dashes, estimate) %>% 
  arrange(factor(window, levels=c('prenatal', 'birth', 'three_year')))

# which edges show both main effects and interaction effects?
#links %>% 
#  left_join(
#    links %>% # count terms per edge
#      group_by(from, to) %>%
#      summarize(n_terms = n_distinct(term), .groups='drop'), 
#    by = c('from', 'to')) %>% 
#  mutate(term = ifelse(n_terms == 2, 'both', term)) %>% 
#  select(-n_terms) %>% 
#  distinct()
```

## Network Visualizations

The first network visualization summarizes the **main effects of chemical mixtures (in solid arrows)**. Edges representing the significant effects were colored based on whether the relationship was positive ('red arrow') or negative ('green arrow').

**Window-specific networks are then presented, with dashed arrows** representing the presence of a significant interaction term (i.e. time window-specific effect), and with main effects being projected again as solid arrows. 

The visualizations are interactive including the following features:  

1. Draggable nodes. 
2. Hovering over a node reveals its name in a pop-up window. 
3. Clicking on a node highlights its connection to other nodes. 
4. Drop-down menu allows for selecting nodes by type.  

```{r}
vis.nodes = nodes %>% filter(prenatal == 1 | all == 1)
vis.links = links %>% filter(window %in% c('prenatal', 'all'))
```


```{r}
# We'll start by adding new node and edge attributes to our dataframes. 
vizNetwork <- function(vis.nodes, vis.links){
  
  vis.nodes$shape  <- "dot"  
  vis.nodes$shadow <- TRUE # Nodes will drop shadow
  vis.nodes$title  <- vis.nodes$variable # Text on click
  vis.nodes$label  <- vis.nodes$variable # Node label
  #vis.nodes$size   <- vis.nodes$audience.size # Node size
  vis.nodes$borderWidth <- 2 # Node border width
  
  colrs <- setNames(c("gray", "tomato", "gold", "dodgerblue", "chartreuse"), 
                    c("Chem. Mix.", "Brain Metabolite", "Cortical Thickness", "Volume", "Subcortical Volume"))
  
  # adapt to variable 
  vis.nodes$color.background <- recode(vis.nodes$group, !!!colrs)
  vis.nodes$color.border <- "black"
  vis.nodes$color.highlight.background <- "orange"
  vis.nodes$color.highlight.border <- "darkred"
  
  edge.colrs <- setNames(c('green', 'red'), c("negative", "positive"))
  vis.links$color <- recode(vis.links$estimate, !!!edge.colrs)    # line color  
  vis.links$width <- ifelse(vis.links$dashes, 1, 2) # line width
  vis.links$arrows <- "to" # arrows: 'from', 'to', or 'middle'
  vis.links$smooth <- TRUE    # should the edges be curved?
  vis.links$shadow <- FALSE    # edge shadow
  vis.links
  
  visnet <- visNetwork(vis.nodes, vis.links, width = 1080) %>% 
    visLayout(randomSeed = 56) %>% 
    visGroups(groupname = "Chem. Mix.", color = 'gray', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Brain Metabolite", color = 'tomato', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Cortical Thickness", color = 'gold', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Volume", color = 'dodgerblue', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Subcortical Volume", color = 'chartreuse', shape = "dot", color.border = 'black') %>% 
    visOptions(highlightNearest = TRUE, selectedBy = 'group') %>% 
      visLegend(addEdges = data.frame(arrows = 'to',
                                      color = c('red', 'green'),
                                      label = c('positive', 'negative')))
    
    
  return(visnet)
}
```

### Main Effect (across all windows)



```{r}
vizNetwork(vis.nodes = nodes %>% filter(all == 1),
           vis.links = links %>% filter(window  == 'all'))
```

<details>
  <summary>Table of Main Effects (click to toggle) </summary>
  
```{r}
options(scipen = 999)
(main_df <- df %>% 
  filter(window == 'all') %>% 
  select(-window) %>% 
  unite(outcome, outcome, outcome.type, sep = ' ') %>% 
  relocate(outcome, effect, .before = 'estimate') %>% 
  arrange(outcome))
```

</details>

### Prenatal 


```{r}
vizNetwork(vis.nodes = nodes %>% filter(prenatal == 1 | all == 1),
           vis.links = links %>% filter(window %in% c('prenatal', 'all')))
```

<details>
  <summary>Table of Interaction Effects (click to toggle) </summary>
  
```{r}
options(scipen = 999)
(prenatal_df <- df %>% 
  filter(window == 'prenatal') %>% 
  select(-window) %>% 
  unite(outcome, outcome, outcome.type, sep = ' ') %>% 
  mutate(outcome = stringr::str_replace(outcome, 'Brain Metabolite', '') %>% 
           stringr::str_replace('Subcortical', '')) %>% 
  relocate(outcome, effect, .before = 'estimate') %>% 
   arrange(outcome))
```

</details>

### Birth

```{r}
vizNetwork(vis.nodes = nodes %>% filter(birth == 1 | all == 1),
           vis.links = links %>% filter(window %in% c('birth', 'all')))
```

<details>
  <summary>Table of Interaction Effects (click to toggle)</summary>
  
```{r}
options(scipen = 999)
(birth_df <- df %>% 
  filter(window == 'birth') %>% 
  select(-window) %>% 
  unite(outcome, outcome, outcome.type, sep = ' ') %>% 
  mutate(outcome = stringr::str_replace(outcome, 'Brain Metabolite', '') %>% 
           stringr::str_replace('Subcortical', '')) %>% 
  relocate(outcome, effect, .before = 'estimate')%>% 
   arrange(outcome))
```

</details>


### Three year

```{r}
vizNetwork(vis.nodes = nodes %>% filter(three_year == 1 | all == 1),
           vis.links = links %>% filter(window %in% c('three_year', 'all')))
```

<details>
  <summary>Table of Interaction Effects (click to toggle) </summary>
  
```{r}
options(scipen = 999)
(three_year_df <- df %>% 
  filter(window == 'three_year') %>% 
  select(-window) %>% 
  unite(outcome, outcome, outcome.type, sep = ' ') %>% 
  mutate(outcome = stringr::str_replace(outcome, 'Brain Metabolite', '') %>% 
           stringr::str_replace('Subcortical', '')) %>% 
  relocate(outcome, effect, .before = 'estimate') %>% 
   arrange(outcome))

```

</details>

## Effect counts across all windows

### By Outcome

```{r}
rbind(main_df, prenatal_df, birth_df, three_year_df) %>% 
  count(outcome, name = 'number_of_effects') %>% 
  arrange(desc(number_of_effects))
```

### By OPE

```{r}
rbind(main_df, prenatal_df, birth_df, three_year_df) %>% 
  count(term, name = 'number_of_effects') %>% 
  arrange(desc(number_of_effects))
```

