---
title: "Network Visualization: Relationship of Chemical Mixtures to Neurotransmitter level, Cortical Thickness and Brain Volumes"
author: "Virgilio Gonzenbach"
date: "`r Sys.Date()`"

output: 
  html_document:
    page_layout: full
    toc: true
    toc_title: 'Contents'
    toc_float: true
    toc_location: left
    df_print: paged
    theme: spacelab
    code_folding: hide
    code_overflow: wrap
    code_tools: true
    fig_width: 10
    fig_height: 7
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      error = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r data_prep}
#| message: false
library(tidyr)
library(dplyr)
library(network)
library(stringr)
library(igraph)
library(visNetwork)

# Read in neurotransmitter data
nt_df <- read.csv(here::here('data/neurotransmitter_interaction.csv'), row.names = NULL) %>% 
  select(-X) %>% 
  pivot_longer(cols = c("mI", "NAA", "Cr", "Cho", "Glu", "Glx"), 
               names_to = 'variable', 
               values_to = 'p.values') %>% 
  tibble::add_column(var.type = 'Neurotransmitter', .after  = 'variable') %>% 
  filter(p.values < 0.1) %>% 
   tibble::add_column(term = 'interaction')

# Read in cortical thickness data
thick_df <- rbind(
  here::here('data/cortical_thickness_interaction.csv') %>% 
    read.csv() %>%
    tibble::add_column(term = 'interaction'),
  here::here('data/cortical_thickness_main.csv') %>% 
    read.csv() %>%
    tibble::add_column(window = 'all', .after = 'OPE') %>% 
    tibble::add_column(term = 'main')
  )  %>% 
  select(-X) %>% 
  rename(variable = brain_region) %>% 
  tibble::add_column(var.type = 'Cortical Thickness', .after  = 'variable')
  
# Read in volumes data
vol_df <- rbind(
  here::here('data/volumes_interaction.csv') %>% 
    read.csv() %>%
    tibble::add_column(term = 'interaction'),
  here::here('data/volumes_main.csv') %>% 
    read.csv() %>%
    tibble::add_column(window = 'all', .after = 'OPE') %>% 
    tibble::add_column(term = 'main')
  ) %>% 
   select(-X) %>% 
  rename(variable = brain_region) %>% 
  tibble::add_column(var.type = 'Volume', .after  = 'variable')
  
# Read in subcortical volumes data
subcort_df <- rbind(
  here::here('data/subcortical_interaction.csv') %>% 
    read.csv() %>%
    tibble::add_column(term = 'interaction'),
  here::here('data/subcortical_main.csv') %>% 
    read.csv() %>%
    tibble::add_column(window = 'all', .after = 'OPE') %>% 
    tibble::add_column(term = 'main')
  ) %>% 
  select(-X) %>% 
  rename(variable = brain_region) %>% 
  tibble::add_column(var.type = 'Subcortical Volume', .after  = 'variable')


# Join all data.frames
df <- rbind(nt_df, 
            thick_df, 
            vol_df,
            subcort_df)
```

The present notebook visualizes statistically significant effects of chemical mixtures to different outcomes: neurotransmitter levels, cortical thickness estimates and cortical and subcortical brain volumes. One network is visualized for each of the following time windows: prenatal, birth, three year.

```{r}
## create nodes 
nodes <- dplyr::union(df %>% select(OPE, var.type, window) %>% rename(variable=OPE), 
             df %>% select(variable, var.type, window)) %>%  
  mutate(var.type = ifelse(variable %in% c("BCEtP", "BDCPP", "DBuP", "DPHP"), 'Chem. Mix.', var.type)) %>% 
  distinct() %>% 
  mutate(value = 1) %>% 
  pivot_wider(names_from = 'window', values_from = 'value', values_fill = 0) %>% 
  mutate(variable = ifelse(var.type == 'Cortical Thickness', paste(variable, '(CT)'), variable)) %>% 
  mutate(variable = ifelse(var.type == 'Volume', paste(variable, '(Vol.)'), variable)) %>% 
  mutate(id = seq_len(nrow(.)), 
         .before = 'variable') %>% 
  rename(group = var.type)
```

```{r}
id_node <- function(node.name){
  nodes %>% 
    filter(variable == node.name) %>% 
    pull(id)
}

links <- df %>% 
  mutate(variable = ifelse(var.type == 'Cortical Thickness', paste(variable, '(CT)'), variable)) %>% 
  mutate(variable = ifelse(var.type == 'Volume', paste(variable, '(Vol.)'), variable)) %>% 
  rename(from = OPE,to = variable) %>% 
  mutate(to = purrr::map_int(to, id_node),
         from = purrr::map_int(from, id_node)) %>% 
  select(from, to, window, term) %>% 
  arrange(factor(window, levels=c('prenatal', 'birth', 'three_year')))

# which edges show both main effects and interaction effects?
#links %>% 
#  left_join(
#    links %>% # count terms per edge
#      group_by(from, to) %>%
#      summarize(n_terms = n_distinct(term), .groups='drop'), 
#    by = c('from', 'to')) %>% 
#  mutate(term = ifelse(n_terms == 2, 'both', term)) %>% 
#  select(-n_terms) %>% 
#  distinct()
```

## Visualizations

### Prenatal

```{r}
#| eval: false
#| echo: false

net <- graph_from_data_frame(d=links %>% 
                               filter(window == 'prenatal'), 
                             vertices=nodes %>% 
                               filter(prenatal == 1), 
                             directed=T) 

colrs <- setNames(c("gray", "tomato", "gold", "dodgerblue", "springgreen2"), unique(nodes$var.type))

V(net)$color <- recode(V(net)$var.type, !!!colrs)
V(net)$label <- NA
V(net)$size <- 20

E(net)$arrow.size <- .4
E(net)$edge.color <- "gray"
```

```{r}
# We'll start by adding new node and edge attributes to our dataframes. 
vizNetwork <- function(vis.nodes, vis.links){
  
  vis.nodes$shape  <- "dot"  
  vis.nodes$shadow <- TRUE # Nodes will drop shadow
  vis.nodes$title  <- vis.nodes$variable # Text on click
  vis.nodes$label  <- vis.nodes$variable # Node label
  #vis.nodes$size   <- vis.nodes$audience.size # Node size
  vis.nodes$borderWidth <- 2 # Node border width
  
  colrs <- setNames(c("gray", "tomato", "gold", "dodgerblue", "chartreuse"), unique(vis.nodes$group))
  
  vis.nodes$color.background <- recode(vis.nodes$group, !!!colrs)
  vis.nodes$color.border <- "black"
  vis.nodes$color.highlight.background <- "orange"
  vis.nodes$color.highlight.border <- "darkred"
  
  edge.colrs <- setNames(c('gray', 'red'), c("interaction", "main"))
  vis.links$color <- recode(vis.links$term, !!!edge.colrs)    # line color  
  #vis.links$width <- 1 # line width
  vis.links$arrows <- "to" # arrows: 'from', 'to', or 'middle'
  vis.links$smooth <- TRUE    # should the edges be curved?
  vis.links$shadow <- FALSE    # edge shadow
  
  visnet <- visNetwork(vis.nodes, vis.links, width = 1080) %>% 
    visLayout(randomSeed = 56) %>% 
    visGroups(groupname = "Chem. Mix.", color = 'gray', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Neurotransmitter", color = 'tomato', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Cortical Thickness", color = 'dodgerblue', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Volume", color = 'gold', shape = "dot", color.border = 'black') %>% 
    visGroups(groupname = "Subcortical Volume", color = 'chartreuse', shape = "dot", color.border = 'black') %>% 
    visLegend(addEdges = data.frame(arrows = 'to', 
                                    color = c('gray', 'red'), 
                                    label = c('interaction', 'main effect'))) %>% 
    visOptions(highlightNearest = TRUE, selectedBy = 'group') 
  return(visnet)
}

vizNetwork(vis.nodes = nodes %>% filter(prenatal == 1 | all == 1),
           vis.links = links %>% filter(window %in% c('prenatal', 'all'))) 
```

### Birth

```{r}
vizNetwork(vis.nodes = nodes %>% filter(birth == 1 | all == 1),
           vis.links = links %>% filter(window %in% c('birth', 'all')))
```


### Three year

```{r}
vizNetwork(vis.nodes = nodes %>% filter(three_year == 1 | all == 1),
           vis.links = links %>% filter(window %in% c('three_year', 'all')))
```

## Data {.tabset}

### Pre-Natal

```{r}
df %>% 
  filter(window == 'prenatal')
```

### Birth

```{r}
df %>% 
  filter(window == 'birth')
```

### 3 year

```{r}
df %>% 
  filter(window == 'three_year')
```
